The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2016-04-10 script/de/chapters/cryptomethods.tex
CT Book ed 12: Corrected some small typos in cryptomethods.tex.

\begin{ctsquote}
    Ein altes Sprichwort, angeblich geprägt von der
    US National Security Agency (NSA), sagt:\\
    ``Angriffe werden immer besser, niemals schlechter.''\\
    ``Attacks always get better; they never get worse.''
\caption[IETF]{IETF\footnotemark}
\end{ctsquote}
\addtocounter{footnote}{0}\footnotetext{%
  \url{http://tools.ietf.org/html/rfc4270}\index{IETF}\index{NSA}
  }

% --------------------------------------------------------------------------
\section[Kryptoanalyse und symmetrische Chiffren für Lehrzwecke]{Kryptoanalyse und symmetrische Chiffren für Lehrzwecke\footnotemark} 
\footnotetext{%
Ein guter Start in die Kryptoanalyse ist das Buch von Mark Stamp \cite{cm:Stamp2007}.\\
Sehr high-level und nur bezogen auf die Analyse symmetrischer Blockchiffren ist
der Artikel von Bruce Schneier \cite{cm:Schneier2000cm}.\\
Einige der Cipher-Challenges bei \glqq MysteryTwister C3\grqq
(\url{http://www.mysterytwisterc3.org}) sind ebenfalls gut für Lehrzwecke
einsetzbar.\index{MTC3}\index{Challenge}\index{Cipher-Challenge}\index{Krypto-Wettbewerb}
}
\index{Kryptoanalyse}

@@

Eine dieser Varianten ist z.B. S-AES (Simplified-AES) von Prof. Ed Schaefer und
seinen Studenten \cite{cm:Musa-etal2003}.%
\footnote{
    Siehe auch die beiden namensgleichen Artikel
    \glqq Devising a Better Way to Teach and Learn the Advanced Encryption Standard\grqq
    unter den Universitäts-News bei\\
    - \url{http://math.scu.edu/~eschaefe/getfile.pdf} \\
    - \url{http://www.scu.edu/cas/research/cryptography.cfm}
}
Eine andere Variante ist Mini-AES \cite{cm:Phan2002}
(siehe das Kapitel~\ref{CM_Sage_Mini-AES} \glqq \nameref{CM_Sage_Mini-AES}\grqq)%
\index{DES}\index{DES!SDES}\index{IDEA}:
\begin{itemize}

@@

\bibitem[Biryukov2009]{cm:Biryukov2009} \index{Biryukov 2009}
	Alex Biryukov, Dmitry Khovratovich, \\
	{\em Related-key Cryptanalysis of the Full AES-192 and AES-256},
	2009 \\
	\url{http://eprint.iacr.org/2009/317}

\bibitem[Coppersmith2002]{cm:Coppersmith2002}  \index{Coppersmith 2002}
       Don Coppersmith, \\
       {\em Re: Impact of Courtois and Pieprzyk results}, \\ 
       19.09.2002, in den \glqq AES Discussion Groups\grqq~ unter \\
       \url{http://aes.nist.gov/aes/}

\bibitem[Courtois2002]{cm:Courtois2002}  \index{Courtois 2002}
       Nicolas Courtois, Josef Pieprzyk, \\
       {\em Cryptanalysis of Block Ciphers with Overdefined Systems
            of Equations}, \\
       received 10 Apr 2002, last revised 9 Nov 2002.\\
       Eine davon verschiedene~\glqq compact version\grqq~des 
       ersten XSL-Angriffs wurde 
       auf der Asiacrypt im Dezember 2002 veröffentlicht. \\
       \url{http://eprint.iacr.org/2002/044}

\bibitem[Ferguson2001]{cm:Ferguson2001}  \index{Ferguson 2001}
       Niels Ferguson, Richard Schroeppel, Doug Whiting, \\
       {\em A simple algebraic representation of Rijndael}, 
       Draft 1. Mai 2001, \\
       \url{http://www.xs4all.nl/~vorpal/pubs/rdalgeq.html}

\bibitem[Haan2008]{cm:Haan2008} \index{Haan 2008} 
       Kristian Laurent Haan, \\
       {\em Advanced Encryption Standard (AES)},
       2008 \\
       \url{http://www.codeplanet.eu/tutorials/cpp/51-advanced-encryption-standard.html}

\bibitem[Hoffman2006]{cm:Hoffman2006} \index{Hoffman 2006} 
       Nick Hoffman, \\
       {\em A SIMPLIFIED IDEA ALGORITHM},
       2006 \\
       \url{http://www.nku.edu/~christensen/simplified%20IDEA%20algorithm.pdf}

@@

\bibitem[Menezes2001]{cm:Menezes2001} \index{Menezes 2001}
       Alfred J. Menezes, Paul C. van Oorschot, Scott A. Vanstone \\
       {\em Handbook of Applied Cryptography}, 
       CRC Press 1997, 5th printing 2001,
       \url{http://www.cacr.math.uwaterloo.ca/hac/} (Errata last update 24.07.2011).

\bibitem[Mironov2006]{cm: Mironov2006} \index{Mironov 2006}
       Ilya Mironov, Lintao Zhang \\
       {\em Applications of SAT Solvers to Cryptanalysis of Hash Functions}, 
       Springer 2006

\bibitem[Musa-etal2003]{cm:Musa-etal2003} \index{Musa, Schaefer, Wedig 2003} 
       Mohammad A. Musa, Edward F. Schaefer, Stephen Wedig, \\
       {\em A simplified AES algorithm and its linear and differential cryptanalyses}, \\
       In: Cryptologia 17 (2), April 2003, Seite~148-177.\\
       Siehe auch:\\
       \url{http://findarticles.com/p/articles/mi_qa3926/is_200304/ai_n9181658}\\
       \url{http://www.rose-hulman.edu/~holden/Preprints/s-aes.pdf}\\
       \url{http://math.scu.edu/~eschaefe/}   Ed Schaefer's Homepage

\bibitem[Nguyen2009]{cm:Nguyen2009} \index{Nguyen 2009} 
       Minh Van Nguyen, \\
       {\em Exploring Cryptography Using the Sage Computer Algebra System}, \\
       Victoria University, 2009, \\
       Siehe SageMath publications \url{http://www.sagemath.org/library-publications.html},\\
       \url{http://www.sagemath.org/files/thesis/nguyen-thesis-2009.pdf},\\
       \url{http://sites.google.com/site/nguyenminh2/honours-thesis-2009.pdf}

\bibitem[Nichols1996]{cm:Nichols1996} \index{Nichols 1996} 
       Randall K. Nichols, \\
       {\em Classical Cryptography Course, Volume 1 and 2}, \\
       Aegean Park Press 1996; \\
       oder in 12 Lektionen online unter \\
       \url{http://www.fortunecity.com/skyscraper/coding/379/lesson1.htm}

\bibitem[Oppliger2011]{cm:Oppliger2011} \index{Oppliger 2011}
       Rolf Oppliger \\
       {\em Contemporary Cryptography, Second Edition},
       Artech House, 2011, \\
       \url{http://books.esecurity.ch/cryptography2e.html}.

@@

\bibitem[Phan2003]{cm:Phan2003} \index{Phan 2003} 
       Raphael Chung-Wei Phan, \\
       {\em Impossible differential cryptanalysis of Mini-AES}, \\
       In: Cryptologia, Oct 2003\\
       \url{http://findarticles.com/p/articles/mi_qa3926/is_200310/ai_n9311721/}

\bibitem[Robshaw2002a]{cm:Robshaw2002a}  \index{Robshaw 2002}
       S.P. Murphy, M.J.B. Robshaw, \\
       {\em Essential Algebraic Structure within the AES},\\
       5. Juni 2002, auf der Crypto 2002,  \\
       \url{http://www.isg.rhul.ac.uk/~mrobshaw/rijndael/rijndael.html}

\bibitem[Robshaw2002b]{cm:Robshaw2002b}  \index{Robshaw 2002}
       S.P. Murphy, M.J.B. Robshaw, \\
       {\em Comments on the Security of the AES and the XSL Technique},\\
       26. September 2002, \\
       \url{http://www.isg.rhul.ac.uk/~mrobshaw/rijndael/rijndael.html}

\bibitem[Schaefer1996]{cm:Schaefer1996} \index{Schaefer 1996} 
       Edward F. Schaefer, \\
       {\em A Simplified Data Encryption Standard Algorithm}, \\
       In: Cryptologia 20 (1), 1996, Seite~77-84

@@

\bibitem[Schneier2000]{cm:Schneier2000cm} \index{Schneier 2000} 
       Bruce Schneier, \\
       {\em A Self-Study Course in Block-Cipher Cryptanalysis}, \\
       Cryptologia, v.24, Jan 2000, pp. 18-34\\
       \url{www.schneier.com/paper-self-study.pdf}

\bibitem[Stallings2006]{cm:Stallings2006} \index{Stallings 2006} 
       William Stallings, \\
       {\em Cryptography and Network Security}, \\
       Prentice Hall 2006,\\
       \url{http://williamstallings.com/}

\bibitem[Stamp2007]{cm:Stamp2007} \index{Stamp 2007} 
       Mark Stamp, Richard M. Low, \\
       {\em Applied Cryptanalysis: Breaking Ciphers in the Real World}, \\
       Wiley-IEEE Press 2007, \\
       \url{http://cs.sjsu.edu/faculty/stamp/crypto/}

>>>>

\begin{ctsquote}
    Ein altes Sprichwort, angeblich geprägt von der
    US National Security Agency (NSA), sagt:\\
    \glqq Angriffe werden immer besser, niemals schlechter.\grqq\\
    ``Attacks always get better; they never get worse.''
\caption[IETF]{IETF\footnotemark}
\end{ctsquote}
\addtocounter{footnote}{0}\footnotetext{%
  \url{http://tools.ietf.org/html/rfc4270}\index{IETF}\index{NSA}
  }

% --------------------------------------------------------------------------
\section[Kryptoanalyse und symmetrische Chiffren für Lehrzwecke]{Kryptoanalyse und symmetrische Chiffren für Lehrzwecke\footnotemark} 
\footnotetext{%
Ein guter Start in die Kryptoanalyse ist das Buch von Mark Stamp \cite{cm:Stamp2007}.\\
Sehr high-level und nur bezogen auf die Analyse symmetrischer Blockchiffren ist
der Artikel von Bruce Schneier \cite{cm:Schneier2000cm}.\\
Einige der Cipher-Challenges bei \glqq MysteryTwister C3\grqq~
(\url{http://www.mysterytwisterc3.org}) sind ebenfalls gut für Lehrzwecke
einsetzbar.\index{MTC3}\index{Challenge}\index{Cipher-Challenge}\index{Krypto-Wettbewerb}
}
\index{Kryptoanalyse}

@@

Eine dieser Varianten ist z.B. S-AES (Simplified-AES) von Prof. Ed Schaefer und
seinen Studenten \cite{cm:Musa-etal2003}.%
\footnote{
    Siehe auch die beiden namensgleichen Artikel
    \glqq Devising a Better Way to Teach and Learn
    the Advanced Encryption Standard\grqq~
    unter den Universitäts-News bei\\
    - \url{http://math.scu.edu/~eschaefe/getfile.pdf} \\
    - \url{http://www.scu.edu/cas/research/cryptography.cfm}
}
Eine andere Variante ist Mini-AES \cite{cm:Phan2002}
(siehe das Kapitel~\ref{CM_Sage_Mini-AES} \glqq \nameref{CM_Sage_Mini-AES}\grqq)%
\index{DES}\index{DES!SDES}\index{IDEA}:
\begin{itemize}

@@

\bibitem[Biryukov2009]{cm:Biryukov2009} \index{Biryukov 2009}
	Alex Biryukov, Dmitry Khovratovich,\\
	{\em Related-key Cryptanalysis of the Full AES-192 and AES-256},
	2009.\\
	\url{http://eprint.iacr.org/2009/317}

\bibitem[Coppersmith2002]{cm:Coppersmith2002}  \index{Coppersmith 2002}
       Don Coppersmith, \\
       {\em Re: Impact of Courtois and Pieprzyk results},\\ 
       19.09.2002, in den \glqq AES Discussion Groups\grqq~unter\\
       \url{http://aes.nist.gov/aes/}

\bibitem[Courtois2002]{cm:Courtois2002}  \index{Courtois 2002}
       Nicolas Courtois, Josef Pieprzyk, \\
       {\em Cryptanalysis of Block Ciphers with Overdefined Systems
            of Equations}, \\
       received 10 Apr 2002, last revised 9 Nov 2002.\\
       Eine davon verschiedene~\glqq compact version\grqq~des ersten
       XSL-Angriffs wurde auf der Asiacrypt im Dezember 2002 veröffentlicht.\\
       \url{http://eprint.iacr.org/2002/044}

\bibitem[Ferguson2001]{cm:Ferguson2001}  \index{Ferguson 2001}
       Niels Ferguson, Richard Schroeppel, Doug Whiting, \\
       {\em A simple algebraic representation of Rijndael}, 
       Draft 1. Mai 2001. \\
       \url{http://www.xs4all.nl/~vorpal/pubs/rdalgeq.html}

\bibitem[Haan2008]{cm:Haan2008} \index{Haan 2008} 
       Kristian Laurent Haan, \\
       {\em Advanced Encryption Standard (AES)},
       2008. \\
       \url{http://www.codeplanet.eu/tutorials/cpp/51-advanced-encryption-standard.html}

\bibitem[Hoffman2006]{cm:Hoffman2006} \index{Hoffman 2006} 
       Nick Hoffman, \\
       {\em A SIMPLIFIED IDEA ALGORITHM},
       2006. \\
       \url{http://www.nku.edu/~christensen/simplified%20IDEA%20algorithm.pdf}

@@

\bibitem[Menezes2001]{cm:Menezes2001} \index{Menezes 2001}
       Alfred J. Menezes, Paul C. van Oorschot, Scott A. Vanstone \\
       {\em Handbook of Applied Cryptography}, 
       CRC Press 1997, 5th printing 2001.\\
       \url{http://www.cacr.math.uwaterloo.ca/hac/} (Errata last update 24.07.2011)

\bibitem[Mironov2006]{cm: Mironov2006} \index{Mironov 2006}
       Ilya Mironov, Lintao Zhang \\
       {\em Applications of SAT Solvers to Cryptanalysis of Hash Functions}, 
       Springer 2006.

\bibitem[Musa-etal2003]{cm:Musa-etal2003} \index{Musa, Schaefer, Wedig 2003} 
       Mohammad A. Musa, Edward F. Schaefer, Stephen Wedig, \\
       {\em A simplified AES algorithm and its linear and differential cryptanalyses}, \\
       In: Cryptologia 17 (2), April 2003, Seite~148-177.\\
       Siehe auch:\\
%       \url{http://findarticles.com/p/articles/mi_qa3926/is_200304/ai_n9181658}\\
       \url{http://www.rose-hulman.edu/~holden/Preprints/s-aes.pdf}\\
       \url{http://math.scu.edu/~eschaefe/}   Ed Schaefer's Homepage

\bibitem[Nguyen2009]{cm:Nguyen2009} \index{Nguyen 2009} 
       Minh Van Nguyen, \\
       {\em Exploring Cryptography Using the Sage Computer Algebra System}, \\
       Victoria University, 2009. \\
       Siehe SageMath publications \url{http://www.sagemath.org/library-publications.html},\\
       \url{http://www.sagemath.org/files/thesis/nguyen-thesis-2009.pdf},\\
       \url{http://sites.google.com/site/nguyenminh2/honours-thesis-2009.pdf}

\bibitem[Nichols1996]{cm:Nichols1996} \index{Nichols 1996} 
       Randall K. Nichols, \\
       {\em Classical Cryptography Course, Volume 1 and 2}, \\
       Aegean Park Press 1996; \\
       oder in 12 Lektionen online unter: \\
       \url{http://www.fortunecity.com/skyscraper/coding/379/lesson1.htm}

\bibitem[Oppliger2011]{cm:Oppliger2011} \index{Oppliger 2011}
       Rolf Oppliger \\
       {\em Contemporary Cryptography, Second Edition},
       Artech House, 2011. \\
       \url{http://books.esecurity.ch/cryptography2e.html}.

@@

\bibitem[Phan2003]{cm:Phan2003} \index{Phan 2003} 
       Raphael Chung-Wei Phan, \\
       {\em Impossible differential cryptanalysis of Mini-AES}, \\
       In: Cryptologia, Oct 2003.\\
       \url{http://findarticles.com/p/articles/mi_qa3926/is_200310/ai_n9311721/}

\bibitem[Robshaw2002a]{cm:Robshaw2002a}  \index{Robshaw 2002}
       S.P. Murphy, M.J.B. Robshaw, \\
       {\em Essential Algebraic Structure within the AES},\\
       5. Juni 2002, auf der Crypto 2002.  \\
       \url{http://www.isg.rhul.ac.uk/~mrobshaw/rijndael/rijndael.html}

\bibitem[Robshaw2002b]{cm:Robshaw2002b}  \index{Robshaw 2002}
       S.P. Murphy, M.J.B. Robshaw, \\
       {\em Comments on the Security of the AES and the XSL Technique},\\
       26. September 2002. \\
       \url{http://www.isg.rhul.ac.uk/~mrobshaw/rijndael/rijndael.html}

\bibitem[Schaefer1996]{cm:Schaefer1996} \index{Schaefer 1996} 
       Edward F. Schaefer, \\
       {\em A Simplified Data Encryption Standard Algorithm}, \\
       In: Cryptologia 20 (1), 1996, Seite~77-84.

@@

\bibitem[Schneier2000]{cm:Schneier2000cm} \index{Schneier 2000} 
       Bruce Schneier, \\
       {\em A Self-Study Course in Block-Cipher Cryptanalysis}, \\
       In: Cryptologia, v.24, Jan 2000, Seite~18-34.\\
       \url{www.schneier.com/paper-self-study.pdf}

\bibitem[Stallings2006]{cm:Stallings2006} \index{Stallings 2006} 
       William Stallings, \\
       {\em Cryptography and Network Security}, \\
       Prentice Hall 2006.\\
       \url{http://williamstallings.com/}

\bibitem[Stamp2007]{cm:Stamp2007} \index{Stamp 2007} 
       Mark Stamp, Richard M. Low, \\
       {\em Applied Cryptanalysis: Breaking Ciphers in the Real World}, \\
       Wiley-IEEE Press 2007. \\
       \url{http://cs.sjsu.edu/faculty/stamp/crypto/}

### 2016-04-10 script/en/chapters/cryptomethods.tex
CT Book ed 12: Corrected some typos and updated some parts in the chapters about bitciphers and ECC.
CT Book ed 12: Corrected some small typos in cryptomethods.tex.

\begin{ctsquote}
    There is an old saying inside the US National Security Agency (NSA):\\
    "Attacks always get better; they never get worse."
\caption[IETF]{IETF\footnotemark}
\end{ctsquote}
\addtocounter{footnote}{0}\footnotetext{%
  \url{http://tools.ietf.org/html/rfc4270}\index{IETF}\index{NSA}
  }

@@

\bibitem[Biryukov2009]{cm:Biryukov2009} \index{Biryukov 2009}
	Alex Biryukov, Dmitry Khovratovich, \\
	{\em Related-key Cryptanalysis of the Full AES-192 and AES-256},
	2009 \\
	\url{http://eprint.iacr.org/2009/317}

@@

\bibitem[Ferguson2001]{cm:Ferguson2001}  \index{Ferguson 2001}
       Niels Ferguson, Richard Schroeppel, Doug Whiting, \\
       {\em A simple algebraic representation of Rijndael}, 
       Draft 2001/05/1, \\
       \url{http://www.xs4all.nl/~vorpal/pubs/rdalgeq.html}

\bibitem[Haan2008]{cm:Haan2008} \index{Haan 2008} 
       Kristian Laurent Haan, \\
       {\em Advanced Encryption Standard (AES)},
       2008, (German article)\\
       \url{http://www.codeplanet.eu/tutorials/cpp/51-advanced-encryption-standard.html}

\bibitem[Hoffman2006]{cm:Hoffman2006} \index{Hoffman 2006} 
       Nick Hoffman, \\
       {\em A SIMPLIFIED IDEA ALGORITHM},
       2006 \\
       \url{http://www.nku.edu/~christensen/simplified%20IDEA%20algorithm.pdf}

@@

\bibitem[Mansoori-etal2007]{cm:Mansoori-etal2007} \index{Mansoori, Bizaki 2007} 
       S. Davod. Mansoori, H. Khaleghei Bizaki, \\
       {\em On the vulnerability of Simplified AES Algorithm Against Linear
       Cryptanalysis}, \\
       In: IJCSNS International Journal of Computer Science and Network
       Security, Vol. 7 No. 7, July 2007, pp~257-263 \\
       See also:\\
       \url{http://paper.ijcsns.org/07_book/200707/20070735.pdf}

@@

\bibitem[Menezes2001]{cm:Menezes2001} \index{Menezes 2001}
       Alfred J. Menezes, Paul C. van Oorschot, Scott A. Vanstone \\
       {\em Handbook of Applied Cryptography}, 
       CRC Press 1997, 5th printing 2001,\\
       \url{http://www.cacr.math.uwaterloo.ca/hac/} (Errata last updated July 24, 2011).

\bibitem[Mironov2006]{cm: Mironov2006} \index{Mironov 2006}
       Ilya Mironov, Lintao Zhang \\
       {\em Applications of SAT Solvers to Cryptanalysis of Hash Functions}, 
       Springer 2006

\bibitem[Musa-etal2003]{cm:Musa-etal2003} \index{Musa, Schaefer, Wedig 2003} 
       Mohammad A. Musa, Edward F. Schaefer, Stephen Wedig, \\
       {\em A simplified AES algorithm and its linear and differential cryptanalyses}, \\
       In: Cryptologia 17 (2), April 2003, pp~148-177.\\
       See also:\\
       \url{http://findarticles.com/p/articles/mi_qa3926/is_200304/ai_n9181658}\\
       \url{http://www.rose-hulman.edu/~holden/Preprints/s-aes.pdf}\\
       \url{http://math.scu.edu/~eschaefe/}   Ed Schaefer's Homepage

\bibitem[Nguyen2009]{cm:Nguyen2009} \index{Nguyen 2009} 
       Minh Van Nguyen, \\
       {\em Exploring Cryptography Using the Sage Computer Algebra System}, \\
       Victoria University, 2009, \\
       See SageMath publications \url{http://www.sagemath.org/library-publications.html},\\
       \url{http://www.sagemath.org/files/thesis/nguyen-thesis-2009.pdf},\\
       \url{http://sites.google.com/site/nguyenminh2/honours-thesis-2009.pdf}

\bibitem[Nichols1996]{cm:Nichols1996} \index{Nichols 1996} 
       Randall K. Nichols, \\
       {\em Classical Cryptography Course, Volume 1 and 2}, \\
       Aegean Park Press 1996;\\
       or in 12 lessons online at \\
       \url{http://www.fortunecity.com/skyscraper/coding/379/lesson1.htm}

\bibitem[Oppliger2011]{cm:Oppliger2011} \index{Oppliger 2011}
       Rolf Oppliger \\
       {\em Contemporary Cryptography, Second Edition},
       Artech House, 2011, \\
       \url{http://books.esecurity.ch/cryptography2e.html}.

@@

\bibitem[Phan2003]{cm:Phan2003} \index{Phan 2003} 
       Raphael Chung-Wei Phan, \\
       {\em Impossible differential cryptanalysis of Mini-AES}, \\
       In: Cryptologia, Oct 2003\\
       \url{http://findarticles.com/p/articles/mi_qa3926/is_200310/ai_n9311721/}

\bibitem[Robshaw2002a]{cm:Robshaw2002a}  \index{Robshaw 2002}
       S.P. Murphy, M.J.B. Robshaw, \\
       {\em Essential Algebraic Structure within the AES}, \\
       June 5, 2002, Crypto 2002,  \\
       \url{http://www.isg.rhul.ac.uk/\~{}mrobshaw/rijndael/rijndael.html}

\bibitem[Robshaw2002b]{cm:Robshaw2002b}  \index{Robshaw 2002}
       S.P. Murphy, M.J.B. Robshaw, \\
       {\em Comments on the Security of the AES and the XSL Technique}, \\
       September 26, 2002, \\
       \url{http://www.isg.rhul.ac.uk/\~{}mrobshaw/rijndael/rijndael.html}

\bibitem[Schaefer1996]{cm:Schaefer1996} \index{Schaefer 1996} 
       Edward F. Schaefer, \\
       {\em A Simplified Data Encryption Standard Algorithm}, \\
       In: Cryptologia 20 (1), 1996, pp~77-84

\bibitem[Schmeh2003]{cm:Schmeh2003}  \index{Schmeh 2003}
       Klaus Schmeh, \\
       {\em Cryptography and Public Key Infrastructures on the Internet},\\ 
       John Wiley \& Sons Ltd., Chichester 2003. \\
       An easy to read book, which also considers practical
       problems such as standardization or real existing software.
       In German, the 5th edition was published in 2013.

@@

\bibitem[Schneier2000]{cm:Schneier2000cm} \index{Schneier 2000} 
       Bruce Schneier, \\
       {\em A Self-Study Course in Block-Cipher Cryptanalysis}, \\
       Cryptologia, v.24, Jan 2000, pp. 18-34\\
       \url{www.schneier.com/paper-self-study.pdf}

\bibitem[Stallings2006]{cm:Stallings2006} \index{Stallings 2006} 
       William Stallings, \\
       {\em Cryptography and Network Security}, \\
       Prentice Hall 2006,\\
       \url{http://williamstallings.com/}

\bibitem[Stamp2007]{cm:Stamp2007} \index{Stamp 2007} 
       Mark Stamp, Richard M. Low, \\
       {\em Applied Cryptanalysis: Breaking Ciphers in the Real World}, \\
       Wiley-IEEE Press 2007, \\
       \url{http://cs.sjsu.edu/faculty/stamp/crypto/}

>>>>

\begin{ctsquote}
    There is an old saying inside the US National Security Agency (NSA):\\
    ``Attacks always get better; they never get worse.''
\caption[IETF]{IETF\footnotemark}
\end{ctsquote}
\addtocounter{footnote}{0}\footnotetext{%
  \url{http://tools.ietf.org/html/rfc4270}\index{IETF}\index{NSA}
  }

@@

\bibitem[Biryukov2009]{cm:Biryukov2009} \index{Biryukov 2009}
	Alex Biryukov, Dmitry Khovratovich, \\
	{\em Related-key Cryptanalysis of the Full AES-192 and AES-256},
	2009. \\
	\url{http://eprint.iacr.org/2009/317}

@@

\bibitem[Ferguson2001]{cm:Ferguson2001}  \index{Ferguson 2001}
       Niels Ferguson, Richard Schroeppel, Doug Whiting, \\
       {\em A simple algebraic representation of Rijndael}, 
       Draft 2001/05/1.\\
       \url{http://www.xs4all.nl/~vorpal/pubs/rdalgeq.html}

\bibitem[Haan2008]{cm:Haan2008} \index{Haan 2008} 
       Kristian Laurent Haan, \\
       {\em Advanced Encryption Standard (AES)},
       2008, German article.\\
       \url{http://www.codeplanet.eu/tutorials/cpp/51-advanced-encryption-standard.html}

\bibitem[Hoffman2006]{cm:Hoffman2006} \index{Hoffman 2006} 
       Nick Hoffman, \\
       {\em A SIMPLIFIED IDEA ALGORITHM},
       2006. \\
       \url{http://www.nku.edu/~christensen/simplified%20IDEA%20algorithm.pdf}

@@

\bibitem[Mansoori-etal2007]{cm:Mansoori-etal2007} \index{Mansoori, Bizaki 2007} 
       S. Davod. Mansoori, H. Khaleghei Bizaki, \\
       {\em On the vulnerability of Simplified AES Algorithm Against Linear
       Cryptanalysis}, \\
       In: IJCSNS International Journal of Computer Science and Network
       Security, Vol. 7 No. 7, July 2007, pp~257-263. \\
       See also:\\
       \url{http://paper.ijcsns.org/07_book/200707/20070735.pdf}

@@

\bibitem[Menezes2001]{cm:Menezes2001} \index{Menezes 2001}
       Alfred J. Menezes, Paul C. van Oorschot, Scott A. Vanstone \\
       {\em Handbook of Applied Cryptography}, 
       CRC Press 1997, 5th printing 2001.\\
       \url{http://www.cacr.math.uwaterloo.ca/hac/} (Errata last updated July 24, 2011).

\bibitem[Mironov2006]{cm: Mironov2006} \index{Mironov 2006}
       Ilya Mironov, Lintao Zhang \\
       {\em Applications of SAT Solvers to Cryptanalysis of Hash Functions}, 
       Springer 2006.

\bibitem[Musa-etal2003]{cm:Musa-etal2003} \index{Musa, Schaefer, Wedig 2003} 
       Mohammad A. Musa, Edward F. Schaefer, Stephen Wedig, \\
       {\em A simplified AES algorithm and its linear and differential cryptanalyses}, \\
       In: Cryptologia 17 (2), April 2003, pp~148-177.\\
       See also:\\
%       \url{http://findarticles.com/p/articles/mi_qa3926/is_200304/ai_n9181658}\\
       \url{http://www.rose-hulman.edu/~holden/Preprints/s-aes.pdf}\\
       \url{http://math.scu.edu/~eschaefe/}   Ed Schaefer's Homepage

\bibitem[Nguyen2009]{cm:Nguyen2009} \index{Nguyen 2009} 
       Minh Van Nguyen, \\
       {\em Exploring Cryptography Using the Sage Computer Algebra System}, \\
       Victoria University, 2009. \\
       See SageMath publications \url{http://www.sagemath.org/library-publications.html},\\
       \url{http://www.sagemath.org/files/thesis/nguyen-thesis-2009.pdf},\\
       \url{http://sites.google.com/site/nguyenminh2/honours-thesis-2009.pdf}

\bibitem[Nichols1996]{cm:Nichols1996} \index{Nichols 1996} 
       Randall K. Nichols, \\
       {\em Classical Cryptography Course, Volume 1 and 2}, \\
       Aegean Park Press 1996;\\
       or in 12 lessons online at: \\
       \url{http://www.fortunecity.com/skyscraper/coding/379/lesson1.htm}

\bibitem[Oppliger2011]{cm:Oppliger2011} \index{Oppliger 2011}
       Rolf Oppliger \\
       {\em Contemporary Cryptography, Second Edition},
       Artech House, 2011.\\
       \url{http://books.esecurity.ch/cryptography2e.html}

@@

\bibitem[Phan2003]{cm:Phan2003} \index{Phan 2003} 
       Raphael Chung-Wei Phan, \\
       {\em Impossible differential cryptanalysis of Mini-AES}, \\
       In: Cryptologia, Oct 2003.\\
       \url{http://findarticles.com/p/articles/mi_qa3926/is_200310/ai_n9311721/}

\bibitem[Robshaw2002a]{cm:Robshaw2002a}  \index{Robshaw 2002}
       S.P. Murphy, M.J.B. Robshaw, \\
       {\em Essential Algebraic Structure within the AES}, \\
       June 5, 2002, Crypto 2002.\\
       \url{http://www.isg.rhul.ac.uk/\~{}mrobshaw/rijndael/rijndael.html}

\bibitem[Robshaw2002b]{cm:Robshaw2002b}  \index{Robshaw 2002}
       S.P. Murphy, M.J.B. Robshaw, \\
       {\em Comments on the Security of the AES and the XSL Technique}, \\
       September 26, 2002.\\
       \url{http://www.isg.rhul.ac.uk/\~{}mrobshaw/rijndael/rijndael.html}

\bibitem[Schaefer1996]{cm:Schaefer1996} \index{Schaefer 1996} 
       Edward F. Schaefer, \\
       {\em A Simplified Data Encryption Standard Algorithm}, \\
       In: Cryptologia 20 (1), 1996, pp~77-84.

\bibitem[Schmeh2003]{cm:Schmeh2003}  \index{Schmeh 2003}
       Klaus Schmeh, \\
       {\em Cryptography and Public Key Infrastructures on the Internet},\\ 
       John Wiley \& Sons Ltd., Chichester 2003. \\
       An easy to read book, which also considers practical
       problems such as standardization or real existing software.
       In German, the 6th edition was published in 2016.

@@

\bibitem[Schneier2000]{cm:Schneier2000cm} \index{Schneier 2000} 
       Bruce Schneier, \\
       {\em A Self-Study Course in Block-Cipher Cryptanalysis}, \\
       Cryptologia, v.24, Jan 2000, pp. 18-34.\\
       \url{www.schneier.com/paper-self-study.pdf}

\bibitem[Stallings2006]{cm:Stallings2006} \index{Stallings 2006} 
       William Stallings, \\
       {\em Cryptography and Network Security}, \\
       Prentice Hall 2006.\\
       \url{http://williamstallings.com/}

\bibitem[Stamp2007]{cm:Stamp2007} \index{Stamp 2007} 
       Mark Stamp, Richard M. Low, \\
       {\em Applied Cryptanalysis: Breaking Ciphers in the Real World}, \\
       Wiley-IEEE Press 2007. \\
       \url{http://cs.sjsu.edu/faculty/stamp/crypto/}
